---
title: 'Example on Archdata package: 02 Creation of the groups compositions'
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
##Replot the dendrogram without problematic elements 
From previous exploration it is obvious that TiO~2~ has an outlier value due to a possible typographic error pointed out previously for the sample GA4.

Dendrogram without TiO~2~ (col nº = 7)

```{r, echo=FALSE}
arch_dendro(df_chem[,-c(7)], printDendro = TRUE, nplot=c(2,3), df_raw = df_raw)
```

##Create the groups

###Create the groups AUTOMATICALLY
Create groups (subset of the main dataset) according to the cluster analysis. k=3 means we are cutting it into theree groups since these are very clear (after removing the TiO2). 

```{r}
#OPTION A

df_chem -> x

# transform the numeric variables to log centered ratio
x.clr <- logcenter.tran(x)


HClust <-  stats::hclust(d = dist(x.clr)^2, method = "cen")
groups <- cutree(HClust, k = 3, h = NULL) #a vector with the names and the group assigned according to the dengroram.

GR_GLO <- subset(df_raw, groups==1) #subset of the whole dataset
GR_NF <- subset(df_raw, groups==2) #subset of the whole dataset
GR_WA <- subset(df_raw, groups==3) #subset of the whole dataset

#cbind(df_raw, groups) -> df_raw

#OPTION B (Get the same by a loop)
  #cutree_1h.dendrogram(my_dend, h = 0.9 ) -> DEND_GROUP

      #for (i in unique(DEND_GROUP))
       #       {
        #      subset(df,DEND_GROUP == i) -> GROUP
        #      cat <- paste("GROUP", i,  sep ="_")
         #     GROUP[,-c(ncol(GROUP))] -> GROUP
          #    assign(cat, GROUP)
           #   print(cat)
              
            #  }
```
###Create the groups MANUALLY (Alternative to previous step)
```{r}
#GR_GLO
##write here the samples to include in the first group
gr_glo_vec <- match(c("GA1","GA2","GA3","GA4","GA5","GB1","GB2","GB3","GB4","GB5","GC1","GC2","GC3","GC4","GD1","GD2","GD3", "GD4","GD5","GE1","GE2","GE3"), df_raw$ID) 

GR_GLO <- subset(df_raw[c(gr_glo_vec),]) #create a subset according to the vector above


#GR_NF
gr_nf_vec <- match(c("C01","CO2","CO3","C04","CO5","C06","C07","C08","C09","C10","C11","C12","C13","C14","GO1","G02"), df_raw$ID) 

GR_NF <- subset(df_raw[c(gr_nf_vec),])


#GR_WA
gr_wa_vec <- match(c("T11","T12","T13","T14","T16","A11","Al2","A16","A18","A26"), df_raw$ID) 
GR_WA <- subset(df_raw[c(gr_wa_vec),])
```






###Check groups
```{r}
groups
```



###MVC + SUM TABLE FOR EACH GROUP 
Introduce here the defined groups

```{r}
group_list <- list(GR_GLO, GR_NF, GR_WA) #include the desired groups
names(group_list) <- c("GR-GLO", "GR-NF", "GR-WA") #include the names with "-" instead of "_" to avoid LaTex problems
rm_vars <- c(1:3,13) #included non desired columns (e.g. negative values, categorical...)
```



###Plot MVC and Summary Tables
In this step for each compositional group, the Matrix of Compositional Variation (MVC), its corresponding dendrogram and the summary tables including: the number of individuals (N), the mean concentration, the standard deviation and the relative standard deviation (RSD). 

The values can be formatted according to the desired significant figures or digits, turning signif = TRUE and indicating the number of desired significant figures (by default signif = 3).

To express values with a fixed decimal number, turn signif = FALSE and indicate the desired number of decimmals (by default decimals = 2).  

```{r warning=FALSE}
j <- 1 #important to run the names in the following code

for (i in group_list){
  names(group_list)[j] -> i_name
  j+1 -> j
  message(i_name)
  arch_evenness(i[,-c(rm_vars)])#
  print(arch_sum_table(i[,-c(rm_vars)], signif = FALSE, decimals = 2))
  print(arch_varmat(i[,-c(rm_vars)]))
  #WP arch_heatmap(i[,-c(rm_vars)])   
  }
```



###Generate LaTex Code for each Summary and MVC table 
```{r}
arch_latex(group_list, landscape = FALSE, MVC = TRUE, rm_vars= rm_vars)
```



###Heatmap of each group
```{r}
message("Heatmap of GR_GLO")
arch_heatmap(GR_GLO[,-c(rm_vars)]) 
message("Heatmap of GR_WA")
arch_heatmap(GR_WA[,-c(rm_vars)])  
message("Heatmap of GR_NF")
arch_heatmap(GR_NF[,-c(rm_vars)]) 
```



# Group Membership Probability 
Using Mahalanobis distance and T^2 Hotelling. 

```{r}
test_sample =  GR_GLO["GA5",] #change for the ANID to be tested
#rm_vars previously defined (the categorical data and undesired variables) 
arch_mahala(group_list, test_sample, rm_vars = rm_vars)
```


`



















