---
title: 'Example on Archdata package: 02 Creation of the groups compositions'
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
##Replot the dendrogram without outliers elements
From previous exploration it is obvious that TiO2 has an outlier value due to a possible typographic error pointed out previously for the sample GA4.

Dendrogram without TiO2 (col nº = 7)

```{r, echo=FALSE}
arch_dendro(df_chem[,-c(7)], printDendro = TRUE, nplot=c(2,3), df_raw = df_raw)
```

##Create the groups

Create groups (subset of the main dataset) according to the cluster analysis. k=3 means we are cutting it into theree groups since these are very clear (after removing the TiO2). 

```{r}
#Watch to perform correctly the Hclust 
groups <- cutree(.HClust, k = 3, h = NULL) #a vector with the names and the group assigned according to the dengroram.

GR_GLO <- subset(df_raw, groups==1) #subset of the whole dataset
GR_NF <- subset(df_raw, groups==2) #subset of the whole dataset
GR_WA <- subset(df_raw, groups==3) #subset of the whole dataset

#cbind(df_raw, groups) -> df_raw
```


###Check groups
```{r}
groups
```



###MVC + SUM TABLE FOR EACH GROUP 
Introduce here the defined groups

```{r}
group_list <- list(GR_GLO, GR_NF, GR_WA) #include the desired groups
names(group_list) <- c("GR-GLO", "GR-NF", "GR-WA") #include the names with "-" instead of "_" to avoid LaTex problems
rm_vars <- c(1:3,13) #included non desired columns (e.g. negative values, categorical...)
```



#Plot MVC and Summary Tables
```{r warning=FALSE}
j <- 1 #important to run the names in the following code

for (i in group_list){
  names(group_list)[j] -> i_name
  j+1 -> j
  message(i_name)
  arch_uniformity(i[,-c(rm_vars)])#
  print(sum_table(i[,-c(rm_vars)]))
  print(arch_varmat(i[,-c(rm_vars)]))
 # WP arch_heatmap(i[,-c(rm_vars)])   
  }
```



###Generate LaTex Code for each Summary and MVC table 
```{r}
arch_latex(group_list, landscape = FALSE, MVC = TRUE, rm_vars= rm_vars)
```
```{r}
summary(df_raw)
```



###Heatmap of each group
```{r}
message("Heatmap of GR_GLO")
arch_heatmap(GR_GLO[,-c(rm_vars)]) 
message("Heatmap of GR_WA")
arch_heatmap(GR_WA[,-c(rm_vars)])  
message("Heatmap of GR_NF")
arch_heatmap(GR_NF[,-c(rm_vars)]) 
```


### Group Membership Probability 
Using Mahalanobis distance and T2 Hotelling. 

```{r}
test_sample =  GR_GLO["GA5",] #change for the ANID to be tested
#rm_vars previously defined (are the categorical data) 
arch_mahala(group_list, test_sample, rm_vars = rm_vars)
```


`



















