---
title: 'Example on Archdata package: 02 Creation of the groups compositions'
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


##Replot the dendrogram without outliers elements
From previous exploration it is obvious that TiO2 has an outlier value due to a possible typographic error pointed out previously for the sample GA4.

Dendrogram without TiO2 (col nยบ = 7

```{r, echo=FALSE}
source('../../R_fun/logcenter.tran.R') #for clr
source('../../R_fun/arch_dendro.R')# for dendrogram
arch_dendro(df_chem[,-c(7)], printDendro = TRUE, nplot=c(2,3), df_raw = df_raw)
```

##Create the groups

Create groups (subset of the main dataset) according to the cluster analysis. k=3 means we are cutting it into theree groups since these are very clear (after removing the TiO2). 

```{r}

#Watch to perform correctly the Hclust 
groups <- cutree(.HClust, k = 3, h = NULL) #a vector with the names and the group assigned according to the dengroram.

GR_GLO <- subset(df_raw, groups==1) #subset of the whole dataset
GR_NF <- subset(df_raw, groups==2) #subset of the whole dataset
GR_WA <- subset(df_raw, groups==3) #subset of the whole dataset

#cbind(df_raw, groups) -> df_raw
```

###Check groups
```{r}
groups
```

##GR_GLO group

```{r}
GR_GLO
```

###Variation Matrix (MVC) plot, MVC dendrogram and CodaDendrogram (GR_GLO)

```{r, echo=FALSE}
source('../../R_fun/entropia02.R')#use in uniformity
source('../../R_fun/classifica.R')#used in uniformtity
source('../../R_fun/arch_uniformity.R')#MVC, and CoDa dendros.
arch_uniformity(GR_GLO[,-c(1:3,13)])#Remove categorical columns
```

###Heatmap

```{r}
arch_heatmap(GR_GLO[,-c(1:3,13)])
```



###Summary table (applied to the GR_GLO dataset)
```{r, echo=FALSE}
source('../../R_fun/sum_table.R')
require(RcmdrMisc)
print(sum_table(GR_GLO[,-c(1:3,13)]))  # This code is external
```


###MVC 
```{r}
arch_varmat(GR_GLO[,-c(1:3,13)]) -> MVC_GR_GLO #save the variation matrix
```



##GR_NF group


```{r}
GR_NF
```

###Variation Matrix (MVC) plot, MVC dendrogram and CodaDendrogram (GR_GLO)
```{r, echo=FALSE}
source('../../R_fun/entropia02.R')#use in uniformity
source('../../R_fun/classifica.R')#used in uniformtity
source('../../R_fun/arch_uniformity.R')#MVC, and CoDa dendros.
arch_uniformity(GR_NF[,-c(1:3,13)])#Remove categorical columns
```

###Heatmap

```{r}
arch_heatmap(GR_NF[,-c(1:3,13)])
```


###Summary table (applied to the GR_NF dataset)
```{r, echo=FALSE}
source('../../R_fun/sum_table.R')
require(RcmdrMisc)
print(sum_table(GR_NF[,-c(1:3,13)]))  # This code is external
```



###MVC
```{r}
arch_varmat(GR_NF[,-c(1:3,13)]) -> MVC_GR_NF  #save the variation matrix
```

##GR_WA group

```{r}
GR_WA
```

###Variation Matrix (MVC) plot, MVC dendrogram and CodaDendrogram (GR_GLO)
```{r, echo=FALSE}
source('../../R_fun/entropia02.R')#use in uniformity
source('../../R_fun/classifica.R')#used in uniformtity
source('../../R_fun/arch_uniformity.R')#MVC, and CoDa dendros.
arch_uniformity(GR_WA[,-c(1:3,13)])#Remove categorical columns
```

###Heatmap

```{r}
arch_heatmap(GR_WA[,-c(1:3,13)])
```


###Summary table (applied to the GR_WA dataset)
```{r, echo=FALSE}
source('../../R_fun/sum_table.R')
require(RcmdrMisc)
print(sum_table(GR_WA[,-c(1:3,13)]))  # This code is external
```

###MVC
```{r}
arch_varmat(GR_WA[,-c(1:3,13)]) -> MVC_GR_WA  #save the variation matrix
```



