---
title: 'Example using RBPottery dataset from CRAN Archdata package: Exploratory analysis of archaeological ceramic compositions'
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

##Summary

The Archdata package is a CRAN package, providing several types of data that are typically used in archaeological research. 

In this parcticular R Script we will be using the RBPottery dataset.

## Dataset Description

The RBPottery which is a a dataset containing results of chemical analyses of 48 specimens of Romano-British pottery from 5 sites in 3 regions. 

A data frame with 48 observations on the following 12 variables:

* ID Sample ID
* Kiln Kiln: Gloucester, Llanedeyrn, Caldicot, Islands Thorns, and Ashley Rails
* Region Region: Gloucester, Wales, and New Forest
* Al2O3 Percentage aluminum trioxide
* Fe2O3 Percentage Iron trioxide
* MgO Percentage magnesium oxide
* CaO Percentage calcium oxide
* Na2O Percentage sodium oxide
* K2O Percentage potassium oxide
* TiO2 Percentage titanium dioxide
* MnO Percentage manganese oxide
* BaO Percantage barium oxide

**Details**:

Results of chemical analyses of 48 specimens of Romano-British pottery published by Tubb, et al.
(1980). The numbers are the percentage metal oxide. "Kiln" indicates at which kiln site the pottery
was found. The kiln sites come from three regions (1=Gloucester, (2=Llanedeyrn, 3=Caldicot),
(4=Islands Thorns, 5=Ashley Rails)). The data were scanned from Table 2.2 in Baxter (2003, p.
21) and preserve three probable typographical errors in the original publication. Those errors are
the values for TiO2 in line 4 (sample GA4), for MnO in line 35 (sample C13), and for K2O in line
36 (sample C14). Versions of these data are also available as Pottery in package car, pottery in
package HSAUR, and Pottery2 in package heplots.

**Source**:

David L. Carlson and Georg Roth (2016). archdata: Example Datasets from Archaeological
Research. R package version 1.1. https://CRAN.R-project.org/package=archdata

**References**

Baxter, M. J. 2003. Statistics in Archaeology. Arnold.

Tubb, A., A. J. Parker, and G. Nickless. 1980. The Analysis of Romano-British Pottery by Atomic
Absorption Spectrophotometry. Archaeometry 22: 153-71.

## Load Libraries
```{r}
  require(archdata) # Load CRAN Archdata package
  require(ggplot2)
  require(plyr)
  source('../../R_fun/entropia02.R')#external code (catalan)
  source('../../R_fun/classifica.R')#external code (catalan)
  source('../../R_fun/arch_uniformity.R') #mofified code (english), it uses functions from entropia02.R and classifica.R
  source('../../R_fun/arch_scattermatrix.R')
  source('../../R_fun/logcenter.tran.R')#for clr 
  source('../../R_fun/arch_dendro.R')# for dendrogram
  source('../../R_fun/arch_PCA.R')
  source('../../R_fun/arch_heatmap.R')
  source('../../R_fun/triangles.R')
```

### Load RBPottery from Archdata pachage
```{r}
  data(RBPottery)
  print(RBPottery)
```

###Declare df_raw and df_chem, and set the row.names
```{r}
  row.names(RBPottery) <- RBPottery[,1]
  df_raw <- RBPottery  #we are calling to the whole dataset
  df_chem <- RBPottery[,-c(1:3)] #ignore all the columns containing categorical data, 9 in this case. 
```

###Check the structure of the new dataset (skip if desired)
```{r}
  str(df_raw)
```
In df_chem only integers (int) or nummerical (num). Num class contain decimmals 
```{r}
  str(df_chem)
```

###Facet plots

Most of the sampless are from Gloucester region (n=22), where only one kiln is present. Following by Wales, were two kilns are found (Llanedeyrn (n=14) and Caldicot), the last with only two samples. Finally there are 5 and 5 individuals in the two kilns from New Forest, distributed in the Islands Thorns Kiln and the Ashley Rails kiln.

```{r}
  y <-df_raw[,1:3]

  .df <- data.frame(y = df_raw$ID, #pie charts
                    s = df_raw$Kiln, #riws 
                    t = df_raw$Region) #columns

  .df <- as.data.frame(with(.df, prop.table(table(y, s, t), margin = 2:3)))

  .plot <-  ggplot(data = .df, aes(x = factor(1), y = Freq, fill = y)) + 
            geom_bar(width = 1, stat = "identity") + 
            scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) + 
            coord_polar(theta = "y") + 
            facet_grid(s ~ t) + 
            xlab(NULL) + 
            labs(fill = "ID") + 
            theme_bw(base_size = 8, base_family = "sans") + 
            theme(panel.spacing = unit(0.3, "lines"), legend.position = "right", 
            axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y 
            = element_blank(), axis.ticks = element_blank())

  print(.plot)
```

###Boxplot with the concentrations

Elemental distribution of the chemical elements
```{r}
  for (i in 1:ncol(df_chem)){
  .datt <- data.frame(REGION=df_raw$Region, i= df_chem[,i], KILN=df_raw$Kiln)
  plot(ggplot(.datt, aes(x= KILN, y = i, fill=REGION)) + geom_boxplot() + ylab(colnames(df_chem)[i]))
  }
```

###Create categories based on the calcium content or any other condition (optional)
```{r}
  df_raw$Calcareous <- c() #Crete the desrired column
  
  "Non-Calcareous (<%6CaO)" ->   df_raw$Calcareous[df_raw$CaO < 6]
  "Calcareous (>%6CaO)" ->       df_raw$Calcareous[df_raw$CaO> 6]
  "High Calcareous (>%20CaO)" -> df_raw$Calcareous[df_raw$CaO > 20]
   
  df_raw$Calcareous <- as.factor(df_raw$Calcareous)  #bring back to factors
```

We might want to have a glance at how is the calcium content of the samples 
```{r}
  select(df_raw,c(CaO,Calcareous)) 
```
##Applying statistics to the dataset

###Variation Matrix (MVC) plot, MVC dendrogram and CodaDendrogram

```{r}
  arch_uniformity(df_chem)
```

###Scatter plot Matrix
```{r}
  #!!select this with the variables to use 
  vars <- c("MnO","CaO","Na2O", "TiO2","BaO", "Al2O3") 
  #!! select the name of the cateogires to plot
  group <- "Region" #e.g "Region", "Kiln"...

  #alr conversion
  alr(df_chem,ivar = 2)-> df_alr
  cbind(df_raw[,1:3],df_alr) -> df_alr

  arch_scatter_matrix(df_raw=df_alr, vars, group, title= "RBPottery")
  #a pdf file is saved in the working directory
```

###Cluster Analysis

Create the dendrograms based on the clr data. Required: define df_chem for numerical df_raw whole dataset (containing categorical data). If desired the emf and pdf can be directly saved to the local folder turning the argument of printDendro = TRUE, by default is false.

nplot: is the vector of plots to be displayed. Chose the index of categorical columns to be displayed. 

```{r}
  arch_dendro(df_chem = df_chem, df_raw = df_raw, printDendro = FALSE, nplot=c(2,3))
```

###Principal Component Analysis 

For the alr transformation, the less variable element is generated in arch_uniformity and saved as .lvar.
Here ".lvar" is read automatically. 
nplot: is the vector of plots to be displayed. Chose the index of categorical columns to be displayed. 
printPCA = TRUE for generating pdf and emf files. 

```{r}
  arch_PCA(df_chem, df_raw =df_raw,lvar=.lvar, nplot=c(2,3)) #nplot is the vector containing categories to be shown
```

###Heatmap (table with the concentrations)
```{r}
  arch_heatmap(df_chem)
```

###Ternary Diagram

In ArchFlow a ternary diagram is included, where the values of CaO, \ch Al2O3 and \ch SiO2 are included in a single plot. However, as for the Tubbs's dataset the \ch SiO2 is not provided there is no chance to generate these kind of plots. See the ternary diagram of the case study instead. 
 

```{r}
  #triangles(df_chem)
```

###Outlier detection 

Through the boxplot functions outliers can be removed. 
In this case, TiO2 is chosen to be removed as a variable in order to asses that the groupping remmains similar and therefore, being able to assingn a group to CA4 sample, which is the one outliying due to the value of TiO2. This value, has been reported as a typographic error before by the ArchData package creator. 


###Final Remarks

```{r}
  print("This is the first step of the ArchFlow workflow")
```










.


