---
title: "Exploratory analysis of archaeological ceramic compositions"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
---

## Description
R-Script for the analysis of archaeological ceramics                
License Script: [GPL](http://www.gnu.org/licenses/gpl-3.0.html)


**Details**:
**Source**:
**References**

Check for any negative data.
```{r}
if (min(df_raw2_0[,-c(1:9)]) < 0) {
  cat("Negative data - terminating.")
}else {
    cat("All the values are positive!")
}
```

##Data import and preprocessing
###Load data
Create a dataframe with the raw data on the csv file. For more details on different formats check [this website](https://rstudio-pubs-static.s3.amazonaws.com/1776_dbaebbdbde8d46e693e5cb60c768ba92.html).
```{r}
df_raw2 <-read.csv("data_NAA.csv", sep=";")
```

###Dataset overview
```{r, echo=FALSE}
df_raw2
```

###Declare df_raw2 and df_chem2, and set the row.names
```{r}
row.names(df_raw2) <- df_raw2[,1]
df_chem2 <- df_raw2[,-c(1:14)] #ignore all the columns containing categorical data, 9 in this case. 
```

###Check the structure of the new dataset (skip if desired)
```{r}
str(df_raw2)
```
In df_chem2 only integers (int) or nummerical (num). Num class contain decimmals 
```{r}
str(df_chem2)
```

###Facet plots

Most of the sampless are from Gloucester region (n=22), where only one kiln is present. Following by Wales, were two kilns are found (Llanedeyrn (n=14) and Caldicot), the last with only two samples. Finally there are 5 and 5 individuals in the two kilns from New Forest, distributed in the Islands Thorns Kiln and the Ashley Rails kiln.

```{r, echo=FALSE}
y <-df_raw2[,1:3]
require("ggplot2")
.df <- data.frame(y = df_raw2$SITE, #pie charts
                  s = df_raw2$REGION, #riws 
                  t = df_raw2$MUNICIPALITY) #columns
.df <- as.data.frame(with(.df, prop.table(table(y, s, t), margin = 2:3)))
.plot <- ggplot(data = .df, aes(x = factor(1), y = Freq, fill = y)) + 
  geom_bar(width = 1, stat = "identity") + 
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) + 
  coord_polar(theta = "y") + 
  facet_grid(s ~ t) + 
  xlab(NULL) + 
  labs(fill = "ID") + 
  theme_bw(base_size = 8, base_family = "sans") + 
  theme(panel.spacing = unit(0.3, "lines"), legend.position = "right", 
  axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y 
  = element_blank(), axis.ticks = element_blank())
print(.plot)
```


###Boxplot with the concentrations

Elemental distribution of the chemical elements
```{r, echo=FALSE}
library(plyr)
for (i in 1:ncol(df_chem2)){
.datt <- data.frame(REGION=df_raw2$REGION, i= df_chem2[,i], SITE=df_raw2$SITE)
plot(ggplot(.datt, aes(x= REGION, y = i, fill=SITE)) + geom_boxplot() + ylab(colnames(df_chem2)[i]))
}
```


###Create categories based on the calcium content or any other condition (optional)
```{r}
df_raw2$Calcareous <- c() #Crete the desrired column
   "Non-Calcareous (<%6CaO)" ->   df_raw2$Calcareous[df_raw2$Ca < 10000]
   "Calcareous (>%6CaO)" ->       df_raw2$Calcareous[df_raw2$Ca > 10000]
   "High Calcareous (>%20CaO)" -> df_raw2$Calcareous[df_raw2$Ca > 100000]
   
df_raw2$Calcareous <- as.factor(df_raw2$Calcareous)  #bring back to factors
```

We might want to have a glance at how is the calcium content of the samples 
```{r}

library(dplyr)
select(df_raw2,c(Ca,Calcareous)) 

```
##Applying statistics to the dataset

###Variation Matrix (MVC) plot, MVC dendrogram and CodaDendrogram

```{r, echo=FALSE}
source('~/Programming/Github/arch_flow/R_fun/entropia02.R')#external code (catalan)
source('~/Programming/Github/arch_flow/R_fun/classifica.R')#external code (catalan)
source('~/Programming/Github/arch_flow/R_fun/arch_uniformity.R') #mofified code (english), it uses functions from entropia02.R and classifica.R
arch_uniformity(df_chem2)
```

###Less variying element (lvar)
```{r}
grep(pattern = "Sm", x = colnames(df_chem2))
```

###Colnames
```{r}
colnames(df_raw2)
```


###Scatter plot Matrix
```{r}
#!!select this with the variables to use 
vars <- c("U","Fe","Mn", "Ti","Hf", "Cs") 
#!! select the name of the cateogires to plot
group <- "REGION" #e.g "Region", "Kiln"...

#alr conversion
library("compositions")
alr(df_chem2, lvar = 22)-> df_alr2
cbind(df_raw2[,2:4],df_alr2) -> df_alr2


source('~/Programming/Github/arch_flow/R_Fun/arch_scattermatrix.R')
arch_scatter_matrix(df_raw=df_alr2, vars, group, title= "NAA data")
#a pdf file is saved in the working directory
```

###Cluster Analysis

Create the dendrograms based on the clr data. Required: define df_chem2 for numerical df_raw2 whole dataset (containing categorical data). If desired the emf and pdf can be directly saved to the local folder turning the argument of printDendro = TRUE, by default is false.

nplot: is the vector of plots to be displayed. Chose the index of categorical columns to be displayed. 


```{r, echo=FALSE}
source('~/Programming/Github/arch_flow/R_fun/logcenter.tran.R')#for clr 
source('~/Programming/Github/arch_flow/R_fun/arch_dendro.R')# for dendrogram
arch_dendro(df_chem2, printDendro = FALSE, nplot=c(2:5), df_raw = df_raw2)
```



###Principal Component Analysis 

For the alr transformation, the less variable element is generated in arch_uniformity and saved as .lvar.
Here ".lvar" is read automatically. 
nplot: is the vector of plots to be displayed. Chose the index of categorical columns to be displayed. 
printPCA = TRUE for generating pdf and emf files. 
```{r, echo=FALSE}
source('~/Programming/Github/arch_flow/R_fun/arch_PCA.R')
arch_PCA(df_chem2, lvar=.lvar, nplot=c(2,3), df_raw2) #nplot is the vector containing categories to be shown
```


###Heatmap (table with the concentrations)
```{r, echo=FALSE}
require(d3heatmap)
source('~/Programming/Github/arch_flow/R_fun/arch_heatmap.R')
arch_heatmap(df_chem2)
```

###Ternary Diagram

In ArchFlow a ternary diagram is included, where the values of CaO, \ch Al2O3 and \ch SiO2 are included in a single plot. However, as for the Tubbs's dataset the \ch SiO2 is not provided there is no chance to generate these kind of plots. See the ternary diagram of the case study instead. 
 
###Convert to oxides:
```{r}
source('~/Programming/Github/arch_flow/R_fun/ox_ppm/to_oxides.R')

to_oxides(df_chem2)
```


```{r}
source('~/Programming/Github/arch_flow/R_fun/triangles.R')
triangles(df_chem2)
```


```{r}
conversion_factors <- read_csv2("~/Programming/Github/arch_flow/R_fun/ox_ppm/conversion_factors.csv", col_names = TRUE)
```





###Without Micaceous

```{r}
df_chemic2 <- df_chem2[-c(which(df_raw2$TYPOLOGY.2== "Micaceous")),]
df_rawmic2 <- df_raw2[-c(which(df_raw2$TYPOLOGY.2== "Micaceous")),]
```

```{r, echo=FALSE}
source('~/Programming/Github/arch_flow/R_fun/arch_PCA.R')
arch_PCA(df_chemic2, lvar=.lvar, nplot=c(2,3), df_rawmic2) #nplot is the vector containing categories to be shown
```
###Withouth honey glazed samples

```{r}
remove_honey <- 
  c(
grep("ORD053", row.names(df_chemic2)),
grep("ORD011", row.names(df_chemic2)),
grep("ORD003", row.names(df_chemic2))
)
```

```{r}
rm_var <- c(
grep("Ca", colnames(df_chemic2)),
grep("Rb", colnames(df_chemic2)),
grep("Ta", colnames(df_chemic2)),
grep("Sb", colnames(df_chemic2)),
grep("K", colnames(df_chemic2)),
grep("As", colnames(df_chemic2)),
grep("Co", colnames(df_chemic2)),
grep("Ca", colnames(df_chemic2)),
grep("Na", colnames(df_chemic2))
)
```

###PCA without honey glazed
```{r, echo=FALSE}
source('~/Programming/Github/arch_flow/R_fun/arch_PCA.R')
arch_PCA(df_chemic2[-c(remove_honey),-(rm_var)], lvar=.lvar, nplot=c(2,3), df_rawmic2[-c(remove_honey),]) #nplot is the vector containing categories to be shown
```
###Preliminary notations:
When removing micaceous and honey glazed, this PCA is obtained.
Log is 2 groups.
Show dendrogram with grouppings and PCA LOG + NAj AND BASQUE POTTERY.
Make groups according to it. 






###Outlier detection 

Through the boxplot functions outliers can be removed. 
In this case, TiO2 is chosen to be removed as a variable in order to asses that the groupping remmains similar and therefore, being able to assingn a group to CA4 sample, which is the one outliying due to the value of TiO2. This value, has been reported as a typographic error before by the ArchData package creator. 



###Final Remarks

This is the first step of the ArchFlow workflow. 








.


